---
title: "Edwin Diaz Pitcher Analysis"
output: html_document
date: "2025-04-13"
---

# Intro

This project will be an analysis into Edwin Diaz across 2022, 2024, and 2025 seasons (so far), looking into his decline in effectiveness and his changing approach. While many metrics would indicate a strong 2024 campaign, he was not nearly as effective as in his phenomenal 2022 season, and his 2025 season has gotten off to a troubling start (NOTE: this is being made with data as of 4/13/2025, 5.2 IP / 6 G into the season for Diaz).

```{r}
library(plyr)
library(dplyr)
library(devtools)
library(DT)
library(ggplot2)
library(ggrepel)
library(ggthemes)
library(gridExtra)
library(janitor)
library(plotly)
library(stringr)
library(tidyr)
library(tidyselect)
library(tidyverse)
library(data.table)
library(reactable)
library(lubridate)
library(ggpubr)
library(paletteer)
```

# Data and Plots

First load in data from the three seasons in question and define strike zone dimensions and making proper adjustments (make it from pitcher's view).

```{r}
diaz <- read_csv("diaz.csv")

heat_colors_interpolated <- colorRampPalette(paletteer_d("RColorBrewer::RdBu", 
                                                         n = 9,
                                                         direction = -1))(16)
left <- -8.5 / 12
right <- 8.5 / 12
bottom <- 18.29 / 12
top <- 44.08 / 12
width <- (right - left) / 3
height <- (top - bottom) / 3

diaz <- diaz %>%
  mutate(plate_x = plate_x * -1)

diaz_22 <- diaz %>%
  filter(game_year == "2022")
diaz_24 <- diaz %>%
  filter(game_year == "2024")
diaz_25 <- diaz %>%
  filter(game_year == "2025")
```

Now, onto the plots. First, I want to visualize Diaz's different pitches. We can see that he only has thrown a fastball and slider over these three years (can see this by doing `unique(diaz$pitch_type)`).

```{r}
#2022
diaz_ff_22 <- diaz_22

freq_ff_22 <- diaz_ff_22 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_ff_22$n)

freq_ff_22$usage_percent <- round((freq_ff_22$n / total)*100, digits = 0)
freq_ff_22 <- subset(freq_ff_22, select = -c(n))

diaz_ff_22 <- merge(diaz_ff_22, freq_ff_22)

diaz_ff_22 <- diaz_ff_22 %>%
  filter(pitch_type == "FF")

ff_total_22 <- diaz_ff_22 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  stat_density2d_filled() +
  scale_fill_manual(values = c(heat_colors_interpolated), aesthetics = c("fill" , "color")) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Fastball"), paste(diaz_ff_22$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

diaz_sl_22 <- diaz_22

freq_sl_22 <- diaz_sl_22 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_sl_22$n)

freq_sl_22$usage_percent <- round((freq_sl_22$n / total)*100, digits = 0)
freq_sl_22 <- subset(freq_sl_22, select = -c(n))

diaz_sl_22 <- merge(diaz_sl_22, freq_sl_22)

diaz_sl_22 <- diaz_sl_22 %>%
  filter(pitch_type == "SL")

sl_total_22 <- diaz_sl_22 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  stat_density2d_filled() +
  scale_fill_manual(values = c(heat_colors_interpolated), aesthetics = c("fill" , "color")) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) + 
  coord_fixed() +
  ggtitle(paste("Slider"), paste(diaz_sl_22$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

#2024
diaz_ff_24 <- diaz_24

freq_ff_24 <- diaz_ff_24 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_ff_24$n)

freq_ff_24$usage_percent <- round((freq_ff_24$n / total)*100, digits = 0)
freq_ff_24 <- subset(freq_ff_24, select = -c(n))

diaz_ff_24 <- merge(diaz_ff_24, freq_ff_24)

diaz_ff_24 <- diaz_ff_24 %>%
  filter(pitch_type == "FF")

ff_total_24 <- diaz_ff_24 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  stat_density2d_filled() +
  scale_fill_manual(values = c(heat_colors_interpolated), aesthetics = c("fill" , "color")) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Fastball"), paste(diaz_ff_24$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

diaz_sl_24 <- diaz_24

freq_sl_24 <- diaz_sl_24 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_sl_24$n)

freq_sl_24$usage_percent <- round((freq_sl_24$n / total)*100, digits = 0)
freq_sl_24 <- subset(freq_sl_24, select = -c(n))

diaz_sl_24 <- merge(diaz_sl_24, freq_sl_24)

diaz_sl_24 <- diaz_sl_24 %>%
  filter(pitch_type == "SL")

sl_total_24 <- diaz_sl_24 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  stat_density2d_filled() +
  scale_fill_manual(values = c(heat_colors_interpolated), aesthetics = c("fill" , "color")) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) + 
  coord_fixed() +
  ggtitle(paste("Slider"), paste(diaz_sl_24$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

#2025
diaz_ff_25 <- diaz_25

freq_ff_25 <- diaz_ff_25 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_ff_25$n)

freq_ff_25$usage_percent <- round((freq_ff_25$n / total)*100, digits = 0)
freq_ff_25 <- subset(freq_ff_25, select = -c(n))

diaz_ff_25 <- merge(diaz_ff_25, freq_ff_25)

diaz_ff_25 <- diaz_ff_25 %>%
  filter(pitch_type == "FF")

ff_total_25 <- diaz_ff_25 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  stat_density2d_filled() +
  scale_fill_manual(values = c(heat_colors_interpolated), aesthetics = c("fill" , "color")) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Fastball"), paste(diaz_ff_25$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

diaz_sl_25 <- diaz_25

freq_sl_25 <- diaz_sl_25 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_sl_25$n)

freq_sl_25$usage_percent <- round((freq_sl_25$n / total)*100, digits = 0)
freq_sl_25 <- subset(freq_sl_25, select = -c(n))

diaz_sl_25 <- merge(diaz_sl_25, freq_sl_25)

diaz_sl_25 <- diaz_sl_25 %>%
  filter(pitch_type == "SL")

sl_total_25 <- diaz_sl_25 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  stat_density2d_filled() +
  scale_fill_manual(values = c(heat_colors_interpolated), aesthetics = c("fill" , "color")) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) + 
  coord_fixed() +
  ggtitle(paste("Slider"), paste(diaz_sl_25$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

#2022 Plot
plot_22 <- ggarrange(ff_total_22,
          sl_total_22, nrow = 1, ncol = 2) %>%
  annotate_figure(plot, top = text_grob("Diaz 2022 (Contour)", 
                                        color = "red", face = "italic", size = 14))
plot_24 <- ggarrange(ff_total_24,
          sl_total_24, nrow = 1, ncol = 2) %>%
  annotate_figure(plot, top = text_grob("Diaz 2024 (Contour)", 
                                        color = "red", face = "italic", size = 14))

plot_25 <- ggarrange(ff_total_25,
          sl_total_25, nrow = 1, ncol = 2) %>%
  annotate_figure(plot, top = text_grob("Diaz 2025 (Contour)", 
                                        color = "red", face = "italic", size = 14))

plot_22
plot_24
plot_25
```

One interesting takeaway from these plots is the differentiation in the pitch usage. Diaz is less reliant on his slider compared to his successful 2022 season, and his fastball has begun to creep up higher in the zone. With the fastball climbing up in the zone, his approach seems to be changing - instead of tunneling effectively in the left half of the zone with fastball and slider, he seems to just be trying to blow fastballs by hitters up and away. I also think that this shift is related to changes in control. Diaz has noticeably been working from behind in the count much more often as of late - from [Baseball Savant](https://baseballsavant.mlb.com/savant-player/edwin-diaz-621242?stats=statcast-r-pitching-mlb&season=2025), Diaz's FirstPitchStrike% by season looks like 71.4% (2022), 63.0% (2024), and 50.0% (2025, through April 12 games). Since this is a big dropoff and can explain the tendency for more fastballs, I'd like to look at how he has approached 0-0 counts and if there's a difference over the years.

```{r}
#set counts variable in new df
diaz_counts <- diaz %>%
  mutate(count = paste0(balls, "-", strikes))
#filter for 0-0 count (what we want just for now)
diaz_00 <- diaz_counts %>%
  filter(count == "0-0")

#2022
diaz_ff_22_00 <- diaz_00 %>%
  filter(game_year == "2022")

freq_ff_22_00 <- diaz_ff_22_00 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_ff_22_00$n)

freq_ff_22_00$usage_percent <- round((freq_ff_22_00$n / total)*100, digits = 0)
freq_ff_22_00 <- subset(freq_ff_22_00, select = -c(n))

diaz_ff_22_00 <- merge(diaz_ff_22_00, freq_ff_22_00)

diaz_ff_22_00 <- diaz_ff_22_00 %>%
  filter(pitch_type == "FF")

ff_total_22_00 <- diaz_ff_22_00 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  geom_point(aes(color = pitch_type), size = 1) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Fastball"), paste(diaz_ff_22_00$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))


diaz_sl_22_00 <- diaz_00 %>%
  filter(game_year == "2022")

freq_sl_22_00 <- diaz_sl_22_00 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_sl_22_00$n)

freq_sl_22_00$usage_percent <- round((freq_sl_22_00$n / total)*100, digits = 0)
freq_sl_22_00 <- subset(freq_sl_22_00, select = -c(n))

diaz_sl_22_00 <- merge(diaz_sl_22_00, freq_sl_22_00)

diaz_sl_22_00 <- diaz_sl_22_00 %>%
  filter(pitch_type == "SL")

sl_total_22_00 <- diaz_sl_22_00 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  geom_point(aes(color = pitch_type), size = 1) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Slider"), paste(diaz_sl_22_00$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

#2024
diaz_ff_24_00 <- diaz_00 %>%
  filter(game_year == "2024")

freq_ff_24_00 <- diaz_ff_24_00 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_ff_24_00$n)

freq_ff_24_00$usage_percent <- round((freq_ff_24_00$n / total)*100, digits = 0)
freq_ff_24_00 <- subset(freq_ff_24_00, select = -c(n))

diaz_ff_24_00 <- merge(diaz_ff_24_00, freq_ff_24_00)

diaz_ff_24_00 <- diaz_ff_24_00 %>%
  filter(pitch_type == "FF")

ff_total_24_00 <- diaz_ff_24_00 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  geom_point(aes(color = pitch_type), size = 1) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Fastball"), paste(diaz_ff_24_00$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))


diaz_sl_24_00 <- diaz_00 %>%
  filter(game_year == "2024")

freq_sl_24_00 <- diaz_sl_24_00 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_sl_24_00$n)

freq_sl_24_00$usage_percent <- round((freq_sl_24_00$n / total)*100, digits = 0)
freq_sl_24_00 <- subset(freq_sl_24_00, select = -c(n))

diaz_sl_24_00 <- merge(diaz_sl_24_00, freq_sl_24_00)

diaz_sl_24_00 <- diaz_sl_24_00 %>%
  filter(pitch_type == "SL")

sl_total_24_00 <- diaz_sl_24_00 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  geom_point(aes(color = pitch_type), size = 1) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Slider"), paste(diaz_sl_24_00$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

#2025
diaz_ff_25_00 <- diaz_00 %>%
  filter(game_year == "2025")

freq_ff_25_00 <- diaz_ff_25_00 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_ff_25_00$n)

freq_ff_25_00$usage_percent <- round((freq_ff_25_00$n / total)*100, digits = 0)
freq_ff_25_00 <- subset(freq_ff_25_00, select = -c(n))

diaz_ff_25_00 <- merge(diaz_ff_25_00, freq_ff_25_00)

diaz_ff_25_00 <- diaz_ff_25_00 %>%
  filter(pitch_type == "FF")

ff_total_25_00 <- diaz_ff_25_00 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  geom_point(aes(color = pitch_type), size = 1) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Fastball"), paste(diaz_ff_25_00$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))


diaz_sl_25_00 <- diaz_00 %>%
  filter(game_year == "2025")

freq_sl_25_00 <- diaz_sl_25_00 %>%
  group_by(pitch_type) %>%
  tally()

total <- sum(freq_sl_25_00$n)

freq_sl_25_00$usage_percent <- round((freq_sl_25_00$n / total)*100, digits = 0)
freq_sl_25_00 <- subset(freq_sl_25_00, select = -c(n))

diaz_sl_25_00 <- merge(diaz_sl_25_00, freq_sl_25_00)

diaz_sl_25_00 <- diaz_sl_25_00 %>%
  filter(pitch_type == "SL")

sl_total_25_00 <- diaz_sl_25_00 %>%
  ggplot(aes(x = plate_x, y = plate_z)) +
  geom_point(aes(color = pitch_type), size = 1) +
  geom_segment(x = left, y = bottom, xend = right, yend = bottom) +
  geom_segment(x = left, y = top, xend = right, yend = top) +
  geom_segment(x = left, y = bottom, xend = left, yend = top) +
  geom_segment(x = right, y = bottom, xend = right, yend = top) +
  geom_segment(x = left, y = (bottom + height), xend = right, yend = (bottom + height)) +
  geom_segment(x = left, y = (top - height), xend = right, yend = (top - height)) +
  geom_segment(x = (left + width), y = bottom, xend = (left + width), yend = top) +
  geom_segment(x = (right - width), y = bottom, xend = (right - width), yend = top) +
  geom_segment(x = left, y = 0, xend = right, yend = 0) +
  geom_segment(x = left, y = 0, xend = left, yend = (4.25/12)) +
  geom_segment(x = left, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = (4.25/12), xend = 0, yend = (8.5/12)) +
  geom_segment(x = right, y = 0, xend = right, yend = (4.25/12)) +
  xlim(-3,3) + ylim(0, 5) +
  coord_fixed() +
  ggtitle(paste("Slider"), paste(diaz_sl_25_00$usage_percent,"%")) +
  theme(legend.position = "none",
        plot.title = element_text(color = "black", size = 15, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color = "black", size = 1.5, fill = NA))

#Plots
plot_22 <- ggarrange(ff_total_22_00,
          sl_total_22_00, nrow = 1, ncol = 2) %>%
  annotate_figure(plot, top = text_grob("Diaz 2022, 0-0", 
                                        color = "red", face = "italic", size = 14))
plot_24 <- ggarrange(ff_total_24_00,
          sl_total_24_00, nrow = 1, ncol = 2) %>%
  annotate_figure(plot, top = text_grob("Diaz 2024, 0-0", 
                                        color = "red", face = "italic", size = 14))

plot_25 <- ggarrange(ff_total_25_00,
          sl_total_25_00, nrow = 1, ncol = 2) %>%
  annotate_figure(plot, top = text_grob("Diaz 2025, 0-0", 
                                        color = "red", face = "italic", size = 14))

plot_22
plot_24
plot_25
```

The trend of increasing fastballs shows up here still, and you can see his fastball gravitating upwards in these 0-0 counts. While 2025 is really tough to judge, the difference between 2022 and 2024 shows that his fastball is less clustered in the zone and his command has shifted upwards, even with the slider. {What it shows}.

Something major that should be focused on is Pitch Tempo, or the time between pitches. Diaz missed 2023 with {injury}, which was the same year that the pitch clock was implemented into MLB games to speed up pitchers, hitters, and overall game time. {Say what the stats show}.

{PLOT?}

{Whiff and chase rates in 0-0, percent ahead/behind in count, usages ahead, etc}
